Automatic Parallelization and Transparent Fault Tolerance\\ (New project paper)

### 2016-04-07 pubs/tfp16/paper.tex
sync

             &                & $|$ & \texttt{case} $e$ \texttt{of} $x$ \texttt{\{}  & Case expression, $n \ge 1$\\
             &                &     & \texttt{ } $\mathit{alt}_1$ \texttt{;} \\
             &                &     & \hspace{0.2in} $\dots$ \texttt{;} \\
             &                &     & \texttt{ } $\mathit{alt}_n$ \texttt{\}} \\
%            &                &     & \texttt{\}} \\
\\
Alternatives & $\mathit{alt}$ & ::= & $C\ x_1\dots x_n$ \texttt{->} $e$            & Pattern match, $n \ge 0$ \\
             &                & $|$ & \texttt{\_ ->} $e$                           & Default \\
\\
Heap objects & $\mathit{obj}$ & ::= &\texttt{FUN} $x_1\dots x_n$ \texttt{->} $e$   & Function definition, arity $=n\ge 1$ \\
             &                & $|$ &\texttt{CON} $C\ a_1\dots a_n$                & Saturated constructor, $n \ge 0$ \\
             &                & $|$ &\texttt{THUNK} $e$                            & Thunk---explicit deferred evaluation \\
             &                & $|$ & $\mathit{PAP}\ f\ a_1\dots a_n$               & Evaluation-time partial application \\
             &                & $|$ & $\mathit{BLACKHOLE}$                         & Evaluation-time black hole \\
\\
Object decl. & $\mathit{odecl}$ &     & $x = \mathit{obj}$                           & Simple binding \\
\\
Constructor defn. & $\mathit{con}$  & ::= & $C\ \mathit{type}_i$ & $i \ge 0$ \\
\\
Datatype defn. &  $\mathit{ddecl}$ & ::= & \texttt{data} [\texttt{unboxed}] & User-defined data type  \\
               &                   &     & $C\ x_i =$                       & $i \ge 0, n > 0$         \\
               &                   &     & $\mathit{con}_1 | \dots |  \mathit{con}_n$ \\

\\
Program      & $\mathit{prog}$ & ::= & $\mathit{(o|d)decl}_1$ & Object and data defns \\
             &                 &     & \texttt{ } $\dots$ \texttt{;} \\
             &                 &     & $\mathit{(o|d)decl}_n$ & 
%Program      & $\mathit{prog}$& ::= & $f_1\ =\ \mathit{obj}_1$ \texttt{;}          & $n \ge 1$, distinguished \texttt{main}\\
%             &                &     & \texttt{ } $\dots$ \texttt{;} \\
%             &                &     & $f_n\ =\ \mathit{obj}_n$

@@

Our incarnation of STG, as shown in Figure~\ref{fig:STGsyntax}, is meant to
evoke both an abstract and concrete syntax.  This presentation is adapted from
Marlow and Peyton Jones' \emph{Eval/Apply} paper~\cite{Marlow:2004}, to which
we refer so many times that we will subsequently refer to it as simply
\emph{EA}. Partial application (\emph{PAP}) and black hole (\emph{BLACKHOLE}),
signifying a \emph{THUNK} being evaluated, are only runtime entities and not
actually part of our STG language.

@@

This then gives three reasonable operational modes that are strictly ordered
in terms of termination properties: non-strict; function strict, in which only
those arguments needed by a function call are evaluated; and application
strict, in which all arguments of an application are evaluated.  The other
dimension we vary is whether or not constructors are strict.  Other dimensions
are possible, for example whether \texttt{let} bindings are strict, but we
restrict ourselves to function/application and constructor strictness.  The
STG machinery makes the choice of evaluation strategy easy to change.
%, both in terms of function application and building a constructor in the
%heap.

@@

For unboxed types code generation and runtime support is currently limited
to built-in single-word (64-bit) or smaller types (e.g., \texttt{Int\#},
\texttt{UInt\#}, \texttt{Float\#}, \texttt{Double\#}) and user-defined
enumerations, e.g.,
\begin{verbatim}
     data unboxed Bool# = False# | True# .
\end{verbatim}

>>>>

             &                & $|$ & \texttt{case} $e$ \texttt{of} $x$ \texttt{\{}  & Case expression, $n \ge 1$\\
             &                &     & \texttt{ } $\mathit{alt}_1$ \texttt{;} \\
             &                &     & \hspace{0.2in} $\dots$ \texttt{;} \\
             &                &     & \texttt{ } $\mathit{alt}_n$ \texttt{\}} \\
%            &                &     & \texttt{\}} \\
\\
Alternatives & $\mathit{alt}$ & ::= & $C\ x_1\dots x_n$ \texttt{->} $e$            & Pattern match, $n \ge 0$ \\
             &                & $|$ & \texttt{\_ ->} $e$                           & Default \\
\\
Heap objects & $\mathit{obj}$ & ::= &\texttt{FUN} $x_1\dots x_n$ \texttt{->} $e$   & Function definition, arity $=n\ge 1$ \\
             &                & $|$ &\texttt{CON} $C\ a_1\dots a_n$                & Saturated constructor, $n \ge 0$ \\
             &                & $|$ &\texttt{THUNK} $e$                            & Thunk---explicit deferred evaluation \\
             &                & $|$ & $\texttt{PAP}\ f\ a_1\dots a_n$               & Evaluation-time partial application \\
             &                & $|$ & $\mathit{BLACKHOLE}$                         & Evaluation-time black hole \\
\\
Object decl. & $\mathit{odecl}$ &     & $x = \mathit{obj}$                           & Simple binding \\
\\
Constructor defn. & $\mathit{con}$  & ::= & $C\ \mathit{type}_i$ & $i \ge 0$ \\
\\
Datatype defn. &  $\mathit{ddecl}$ & ::= & \texttt{data} [\texttt{unboxed}] & User-defined data type  \\
               &                   &     & $C\ x_i =$                       & $i \ge 0, n > 0$         \\
               &                   &     & $\mathit{con}_1 | \dots |  \mathit{con}_n$ \\

\\
Program      & $\mathit{prog}$ & ::= & $\mathit{(o|d)decl}_1$ \texttt{;}            & Object and data defns \\
             &                 &     & \texttt{ } $\dots$ \texttt{;} \\
             &                 &     & $\mathit{(o|d)decl}_n$ & 
%Program      & $\mathit{prog}$& ::= & $f_1\ =\ \mathit{obj}_1$ \texttt{;}          & $n \ge 1$, distinguished \texttt{main}\\
%             &                &     & \texttt{ } $\dots$ \texttt{;} \\
%             &                &     & $f_n\ =\ \mathit{obj}_n$

@@

Our incarnation of STG, as shown in Figure~\ref{fig:STGsyntax}, is meant to
evoke both an abstract and concrete syntax.  This presentation is adapted from
Marlow and Peyton Jones' \emph{Eval/Apply} paper~\cite{Marlow:2004}, to which
we refer so many times that we will subsequently refer to it as simply
\emph{EA}.  Black hole (\emph{BLACKHOLE}), signifying a \emph{THUNK} being
evaluated, is only a runtime entity and not actually part of our STG language.

@@

This then gives three reasonable operational modes which are strictly ordered
in terms of termination properties: non-strict; function strict, in which only
those arguments needed by a function call are evaluated; and application
strict, in which all arguments of an application are evaluated.  The other
dimension we vary is whether or not constructors are strict.  Other dimensions
are possible, for example whether \texttt{let} bindings are strict, but we
restrict ourselves to function/application and constructor strictness.  The
STG machinery makes the choice of evaluation strategy easy to change, though
changing code generation can be more efficient by avoiding the unnecessary
creation of \emph{THUNKS} as discussed in Appendix~A.

@@

While we can parse and type-check more general unboxed types, code generation
and runtime support is currently limited to built-in single-word (64-bit) or
smaller types (e.g., \texttt{Int\#}, \texttt{UInt\#}, \texttt{Float\#},
\texttt{Double\#}) and user-defined enumerations (simple sums), e.g.,
\texttt{data unboxed Bool\# = False\# | True\#}.
%\begin{verbatim}
%     data unboxed Bool# = False# | True# .
%\end{verbatim}
